---
title: "COLA RR analysis plan"
output: html_notebook
---
Started 28th August 2020  , DVM Bishop  

Analysis script illustrated with simulated data.  
For each part of analysis, we first show the text from the RR, and then illustrate the analysis with a simulated dataset.

```{r Rpackage_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggstatsplot")
library(ggstatsplot)
library(tidyverse)
library(MASS)
library(MBESS)
library(nlme)
library(semPower)
library(semTools)
library(bookdown)
library(lavaan)
library(semPlot)
# library(knitr)
library(flextable)
library(officer)
library(corrr) #added by DB for easy correlations
library(plyr)
library(qpcR) #used in Kievit script
library(kableExtra)
library(ggpubr)
options(scipen=999)
```


## Simulated data
To test our analysis, we create a simulated dataset, which is roughly based on what we know about distributions of the LIs based on prior research.  

A dataset is simulated for 200 Lhanders and 200 Rhanders on three online laterality tasks and six fTCD tasks. The simulated data are LIs - i.e. these are already computed at this stage.  

```{r simdata}
set.seed(3)
#3 online tasks referred to by name: P1-P3 are production fTCD tasks, R1:R3 are receptive.
mytasks<-c('rhyme','dichotic','ovp','P1','P2','P3','R1','R2','R3')
#We start assuming rhyme will pattern with P tasks, and dichotic and ovp with R tasks

mymeans <- c(.6,.5,.4,1.5,1.5,1.2,.5,.5,.3) #starting values for mean LIs (some cases will be flipped so actual means will be less than this) - some random task effects incorporated
myN <- 400

mySigma <- matrix(c(10,3,3,6,6,6,3,3,3,
                    3,10,6,3,3,3,6,6,6,
                    3,3,10,3,3.,3,6,6,6,
                    6,3,3,10,6,6,3,3,3,
                    6,3,3,6, 10,6,3,3,3,
                    6,3,3,6,6,10,3,3,3,
                    3,6,6,3,3,3,10,6,6,
                    3,6,6,3,3,3,6,10,6,
                    3,6,6,3,3,3,6,6,10),nrow=9)



mySigma<-mySigma/10  #much easier to debug without decimals!

mydat <- mvrnorm(myN,mymeans,mySigma)
Lhand <- c(rep(0,200),rep(1,200))
mydatF <-data.frame(cbind(Lhand,mydat))
colnames(mydatF)[2:10]<-mytasks

#now flip polarity for proportion of individuals.

#Define proportion who flip - NB this is not same as % R-lateralised because initial values include negatives, so some will flip from R to L!

#Do separate flip for production and reception variables, to induce dissociation in some
pL <- .35  #proportion of L handers who flip
pR <- .1 #proportion of R handers who flip
pvars <- c(1,4,5,6)+1  #column numbers for production tasks
rvars <- c(2,3,7,8,9)+1 #column numbers for receptive tasks

for (i in 1:myN){
  myp <- pR #proportion who will flip
  if(i>199){myp <- pL} #use Lhander proportion for rows 200+
  if(runif(1)<myp) {mydatF[i,pvars]<-(-mydatF[i,pvars])} #check p against one random uniform for production
   if(runif(1)<myp) {mydatF[i,rvars]<-(-mydatF[i,rvars])}#check p against another random uniform for receptive
  }
print('Correlation check: whole sample')
correlate(mydatF[,2:10])
print('L handers only')
correlate(mydatF[200:400,2:10])
print('R handers only')
correlate(mydatF[1:200,2:10])

mydatF$id<-as.factor(1:400)
mydatF$Lhand<-as.factor(mydatF$Lhand)

```
## Step 0. 
__Basic descriptives__  
Distributions of data will be visualised in scatterplots, showing pairwise associations between the three LI measures, with handedness colour-coded. As well as providing a table with means and SDs of all variables, we will compute split-half reliability for the online battery measures by correlating LIs obtained with alternate blocks of items for each task, and for the Doppler measures from alternate trials. Spearman correlations will be used to avoid inflated reliability from unduly influential datapoints.


## Step 1. 
__Prediction: On the online battery, dichotic and visual-half-field tasks will show significant left-brain lateralisation at the group level, with this effect being stronger in right- than left-handers.__    

To examine lateralisation at the brain level for each task will conduct the following analyses. Following Parker et al. (2020), we will conduct Shapiro-Wilk tests to check whether the distributions are normal for each measure. If distributions are normal, we will conduct a one-sample t-test to assess whether LIs at the group level are different from zero; otherwise a a Wilcoxon one-sample v test will be used.


## Test for normality

```{r shapirowilk}


  thisdat<-mydatF

  for (t in 2:10){ #each test
    s<-shapiro.test(thisdat[,t])$p.value
    print(paste0(colnames(thisdat)[t],' Normality test, p-value: ',round(s,3)))
    }


```
Note that, with this method of simulation, whether or not non-normality detected will depend on the initial mean LI specified for the variable. Flipping data from a high mean will induce bimodality.
 

## Test whether population mean is different from zero  

```{r onesamplet}


  for (t in 2:10){ #each test
    x<-t.test(thisdat[,t])
    print(paste0(colnames(thisdat)[t],': mean = ',round(x$estimate,3),', t=', round(x$statistic,3),', p-value = ',round(x$p.value,4)))
    }


```

Check distributions comparing L and R handers

```{r densfunction}
# Density plot 
densplotfunc<-function(myfile,mymain){

p <- ggplot(myfile, aes(x=thiscol,color=Lhand)) + 
    geom_density(alpha=.4)+
  ggtitle(mymain)+ # for the main title
xlab('LI') # for the x axis label


# Add mean lines

mu <- ddply(myfile, "Lhand", summarise, grp.mean=mean(thiscol))
p <- p+geom_vline(data=mu, aes(xintercept=grp.mean,color=Lhand), linetype="dashed")
return(p)
}
```

```{r doplots}
#run function for each measure

thisplot <-0

for (col in 2:4){ #do for each online variable
  thisplot<-thisplot+1
  myfile2<-thisdat[,c(1,col)]
  mymain<-colnames(thisdat)[col]
  colnames(myfile2)[2]<-'thiscol'
  p<-densplotfunc(myfile2,mymain)
  t<-t.test(myfile2$thiscol~myfile2$Lhand)
  mymeans<-round(by(myfile2$thiscol,myfile2$Lhand,mean),3)
  mysds<-round(by(myfile2$thiscol,myfile2$Lhand,sd),3)
  print(paste0(mymain,": t=",round(t$statistic,2),",p = ",round(t$p.value,3)))
  print(paste0('Mean Rhander = ',mymeans[1],' SD = ',mysds[1], '.  Mean Lhander = ',mymeans[2],' SD = ',mysds[2]))
#  ggarrange(myg1,myg2,myg3,myg4,myg5,myg6,myg7,myg8, ncol = 2, nrow = 4,common.legend=T,legend="bottom")
# ggsave(
#   "all.raw.png",
#   width = 5, height = 8,
#   dpi = 300
  plotname<-paste0('scatter',thisplot,'.png')
  ggsave(plotname)
  print(p)
}

```

The main analysis of the behavioural laterality indices will follow the approach used by Bruckert et al (2020), using Multilevel Linear Modelling (MLM) to assess the influence of task and handedness on lateralisation strength (LI). This method allows us to test whether the heterogeneity of variance is comparable in left- and right-handers, as well as quantifying main effects of task and handedness. MLM will be conducted within R (R Core Team, 2019) using the nlme package (version 3.1-148; Pinheiro et al., 2020). Following Bruckert et al. two models will be fitted to the data. Model 1 will assume homogeneous variance model assumptions (i.e. variances between subjects and groups are equal for the two handedness groups). Model 2 will assume heterogeneous between-subject variance by group (i.e. different variances for the left and right handed group). Model fit will be assessed using a likelihood ratio test comparing Model 1 and Model 2. Coefficients for the fixed and random effects will be reported for the best fitting model. Both models will include fixed effects of handedness, task and their interaction: nlme::lme(fixed = LI ~ handedness*task, random = list(id = pdDiag(form= ~ 0 + handedness))).  

We will adopt the following contrasts. For handedness, the contr.sum function from the base stats package will be used to implement summed-to-zero contrasts. This will correspond to a main effect of handedness. Helmert contrasts will be used for the task variable which compares each categorical level to the mean of the subsequent levels.


## Multilevel linear model
This is run on the online measures
```{r MLL}
  thisdat<-mydatF[,c(1:4,11)] #handedness, online measures, ID

#First convert files to long form
longdat <- thisdat %>% gather("task", "LI", -c(id,Lhand))

longdat$task<-as.factor(longdat$task)
longdat$hand<-as.factor(longdat$Lhand)


# Homogeneous (base model)
mod0<-lme(fixed=LI~hand*task, 
          random=list(id=pdSymm(form=~1)),
          data=longdat, 
          method="REML",contrasts=c(hand='contr.sum',task='contr.helmert'))

print(summary(mod0))

# Heterogeneous (alternative model)

mod1<-lme(fixed=LI ~ hand*task, 
          random=list(id=pdDiag(form= ~ 0 + hand)),
          data=longdat,
          method="REML",contrasts=c(hand='contr.sum',task='contr.helmert'))

sum.mod1<-summary(mod1)
print(summary(mod1))

#Likelihood ratio test

LRtest <- anova(mod0,mod1)
print('Homogeneous vs heterogeneous model')
print(LRtest)

```

## Step 2
__2. The pattern of correlation between laterality indices from online measures will reflect the extent to which they involve implicit speech production, rather than whether they involve spoken or written language. Thus we anticipate dissociation between the rhyme judgement task and the other two measures (dichotic listening and OVP task), which is not accountable for in terms of low reliability of measures.__ 

For prediction 2 we focus on predicting patterns of covariance between the three online tasks, as we do not have sufficient indicators for modeling with latent variables. We prespecify four possible covariance structures, which are then compared using AIC weights. This is a subset of SEM that does not include any latent variables or directional paths. It allows us to constrain particular covariance patterns and report the 'best' model according to AIC weights.

The four models are:  
Model A, where all LIs are intercorrelated to a similar degree
Model B1,  where LIs for the two receptive language measures (dichotic listening and optimal viewing task) are intercorrelated, but independent of rhyme detection
Model B2, where LIs for the two tasks involving visual presentation and written language (optimal viewing and rhyme detection) are intercorrelated, and independent of the auditory task, dichotic listening.
Model C, where all LIs are independent of one another.

Models B1 and B2 are mathematically equivalent, differing only in the specific variables that are correlated. On a priori grounds, we favour model B1, which is compatible with our overall 2-factor model of language lateralisation. We mention B2, however, as this pattern of correlation might occur if the laterality index was dependent more on mode of presentation (visual vs auditory) than on task demands. In the section below on sample size justification, we outline the syntax used for this method and show that our sample size is adequate to distinguish between models.  Note that where we specify tests as correlated in a model, the extent of correlation will be constrained by test reliability. If the reliability of any of our measures is lower than .5, this would limit the interpretation of the models. 

```{r aicmodels}
#Specify models

  #Model A all correlated
    modelA<-"
    rhyme~~dichotic
    rhyme~~ovp
    dichotic~~ovp
    "
    #Model B 2 correlated (ovp and dichotic)
    modelB<-"
    rhyme~~0*dichotic
    rhyme~~0*ovp
    dichotic~~ovp"
    
    #model C all independent
    modelC<-"
    rhyme~~0*dichotic
    rhyme~~0*ovp
    dichotic~~0*ovp"

    #Model BB 2 correlated (ovp and rhyme)
    modelBB<-"
    rhyme~~0*dichotic
    rhyme~~ovp
    dichotic~~0*ovp"
    
```    

```{r runAIC}
AICdf <- data.frame(matrix(NA,nrow=3,ncol=8))
colnames(AICdf)<-c('dataset','hand','modelA','modelB','modelC','modelBB','bestfit','chi.p.vs.C')
handgps <- c('All','L.only','R.only')
thisrow<-0

thisdatall<-mydatF
thisdat<-thisdatall
  for (h in 1:3){ #handedness grouping - all, L only, R only}
    
    
     if(h==2){thisdat<-thisdatall[201:400,]} #left-handers only
    if(h==3){thisdat<-thisdatall[1:200,]} #right-handers only  

    thisrow<-thisrow+1
 
    fitA <- cfa(modelA, data=thisdat) #
    fitB <- cfa(modelB, data=thisdat)
    fitC <- cfa(modelC, data=thisdat)
    fitBB <- cfa(modelBB,data=thisdat)
    
   aic_vector1<- c(fitMeasures(fitA, "aic"),fitMeasures(fitB, "aic"),fitMeasures(fitC, "aic"),fitMeasures(fitBB,"aic"))
    
    AICdf[thisrow,3:6]<-round(akaike.weights(aic_vector1)$weights,3)
   bestfit <- which(AICdf[thisrow,3:6]==max(AICdf[thisrow,3:6]))
     AICdf$bestfit[thisrow] <- LETTERS[bestfit]
     AICdf$hand[thisrow]<-handgps[h]
     AICdf$dataset[thisrow] <- 'mydatF'

    
 if(bestfit<3)
 {
   compfit<-fitA
   if(bestfit==2){compfit<-fitB}
#If true model is not C, can it be distinguished from model C (all independent)
    AICdf$chi.p.vs.C[thisrow] <- round(anova(compfit,fitC)$`Pr(>Chisq)`[2],3)
 }
 }#end of h loop



flextable(AICdf)
```


# FTCD measures  

## Step 3
__The data will fit a model where 'language generation' tasks cluster together on one factor, and 'receptive language’ tasks on a second factor.__  

It is predicted that factors will be correlated, but the fit of a 2-factor model will be superior to a single-factor model where all LIs load on a common factor.
 
The analysis conducted by Woodhead et al (2019, 2020) used an exploratory bifactor model in which each task could load on each of two factors. Because we had two measures for each task (from test and retest sessions), this exploratory approach was adequately powered. The current study will use confirmatory factor analysis, using a prespecified two-factor model which constrains which indicators can load on two factors. This will be compared to a unitary model, in which all tasks load on a single factor.
 
If the fit of the two-factor model is poor, we will divide the sample into two random halves, before proceeding to drop or add paths from the model in Figure 4 to improve fit. When the optimal model has been identified, it will then be tested in confirmatory factor analysis using the hold-out portion of the data

Note that we now have only 100 participants per handedness group, so we drop half the sample for a more realistic simulation.  

```{r modelfit}
dopdat <- mydatF[c(1:100,201:300),-c(2:4)]

model.h1 <- '
f1 =~ P1 + P2 + P3
f2 =~ R1 + R2 + R3

f1 ~~ 1*f2  #single factor model, f1 and f2 constrained to covariance of one

' 
model.h2 <- '
f1 =~ P1 + P2 + P3
f2 =~ R1 + R2 + R3  #2 factor model: no constraint on covariance

'

fit1 <- cfa(model.h1, data=dopdat)
fit2 <- cfa(model.h2, data=dopdat)
sfit1 <-summary(fit1,fit.measures=TRUE)
sfit2 <- summary(fit2,fit.measures=TRUE)

anova(fit1,fit2)

```


## Step 4. 
__The pattern of variances and covariances will differ for right-handers and left-handers, with better model fit being obtained when separate 2-factor models are estimated for left- vs right-handers.__
This will be tested using the optimal model obtained from step 3.  Here this is the 2-factor model as specified above, which matched the process for simulating the data.  
The measurement invariance command tests a series of models which progressively constrain the equality of parameters between left- and right-handed groups.

```{r testLRdiffmodel}
dopdat<-dopdat[,c(8,1:7)]
levels(dopdat$Lhand)<-c('Right','Left')

fit2a <- cfa(model.h2, data=dopdat,group="Lhand")

sfit2a <- summary(fit2a,fit.measures=TRUE)

measurementInvariance(model=model.h2, data=dopdat,group="Lhand")

```

## Step 5
__On categorical analysis, individuals who depart from left-brained laterality on one or more tasks will be more likely to be left-handed than those who are consistently left-lateralised.__

Prediction 5 moves to categorical analysis. A simple approach is to dichotomise laterality at a cutoff of zero for each task, and then perform a chi square analysis to test for association with handedness.  For 6 measures, we adopt a Bonferroni-corrected alpha level of .05/6 = .008. 

```{r chisq.cats}
#initialise catx variables: 1 for +ve and - for negative
dopdat$catx1<-1
dopdat$catx2<-1
dopdat$catx3<-1
dopdat$catx4<-1
dopdat$catx5<-1
dopdat$catx6<-1

chitab<-data.frame(matrix(NA,ncol=11,nrow=6))
for (mycol in 1: 6){
  w<-which(dopdat[,(mycol+2)]<0) #find values below zero
  dopdat[w,(mycol+8)]<-0 #recode catx value to zero
  t<-table(dopdat[,(mycol+8)],dopdat$Lhand)
  tp<-round(prop.table(t,2),2) #proportions by handedness
  chitab[mycol,1]<-colnames(dopdat)[(mycol+2)]
  chitab[mycol,c(2,4,6,8)]<-t
  chitab[mycol,c(3,5,7,9)]<-tp
  chitab[mycol,10] <- chisq.test(t)$statistic
  chitab[mycol,11] <- chisq.test(t)$p.value
}
colnames(chitab)<-c('Measure','Rhand.Rbrain','RH.Rb%','Rhand.Lbrain','RH.Lb%','Lhand.Rbrain','LH.Rb%','Lhand.Lbrain','LH.Lb%','chisq','p')
flextable(chitab)

```

# Relationship between fTCD and behavioural laterality indices  
## Step 6
__The laterality profile obtained with the online language battery will be significantly associated with the profile seen with direct measure of cerebral blood flow using fTCD, with laterality on dichotic listening and OVP relating more strongly to receptive language tasks, and rhyme judgement to language generation tasks.__

```{r redoCFA}

alldat200 <- mydatF[c(1:100,201:300),]

model.h1 <- '
f1 =~ P1 + P2 + P3 + rhyme
f2 =~ R1 + R2 + R3 + dichotic + ovp

f1 ~~ 1*f2  #single factor model, f1 and f2 constrained to covariance of one

' 
model.h2 <- '
f1 =~ P1 + P2 + P3 + rhyme
f2 =~ R1 + R2 + R3 + dichotic + ovp  #2 factor model: no constraint on covariance

'

fit1 <- cfa(model.h1, data=alldat200)
fit2 <- cfa(model.h2, data=alldat200)
sfit1 <-summary(fit1,fit.measures=TRUE)
sfit2 <- summary(fit2,fit.measures=TRUE)

anova(fit1,fit2)

```

Model 2 gives perfect fit. This is not surprising, since we simulated data assuming model 2. 

## Session information

```{r sessinf}
sessionInfo()
```
